---
title: "Initial Korean covid 19"
author: "davy d"
date: "spring 2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

# Introduction

$H_0$ : There is no difference in death rate based on sex.

$H_A$ : There is a difference in death rate based on sex. 

## METHODS
Variables for this study are sex and result.
We want to know if result depends on Sex result is a categorical variable. Sex is also a categorical variable. So we use the cat~cat helper file. 



## Graphical results - use ggplot commands to do this  

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

This chart shows that males are more likely to die witht ehaverage being 60 percent and female average being below 40 perfcent.  





## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
The table shows the average male deaths are higher than female deaths at 64 deaths versus female deaths at 36. 

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi-squared test p-value is 0.01 which is lesser than 0.05 which means we can effectively reject the null hypothesis this means there is significant difference in death rate based on sex.


P-value =0.0009 on the fisher exact test shows that we can reject the null hypothesis because the value is less than 0.05.

The odds ratios is 2.3 to 1 this means that there is higher odds to prove that men do have a larger death rate than women do. 

## Conclusions

The original question is there a difference in Korean data death rate based on sex? With the numerical and graphical results we are able to come to the conclusion there is a difference between sex and death rate, men do have a higher mortality rate and we where able to prove that within in our boxplot, numerical graphic and lastly our chi-squared test p-value = 0.01 which is less than 0.05 which gives us helps us reject the null hypothesis and prove there is a difference in sex in and death rate.